Random Forests

### 2023-11-12 src/la_RF.Rmd
rf

- [x] read more about trainColntrol
- [ ] other variables, LA Classification, check regressions, 
- [x] N/S
- [x] reproducible
- [ ] n vs. log(n/n.lag), %>% mutate(n.lag = n.lag + 0.00001, n = n + 0.00001, way to bring S?
- [1] center or scale with preProcess??
- [ ] direction of relationship, or cor()
- [x] importance over time
- [2] spatial sampling
- [ ] OA

@@

for (t in years){
  tplus1 <- t + 1
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
                     data = df %>% filter(year==t | year==tplus1), #%>% 
                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
                     trControl = tc,
                     method = "rf", #"rpart", # rf lm
                     na.action = na.omit,
                     importance = TRUE)
  assign(paste0("model.all",as.character(t + 1)), model.all)
}

@@

>>>>

- [x] read more about trainColntrol
- [ ] other variables, LA Classification, check regressions, 
- [x] N/S
- [x] reproducible
- [ ] n vs. log(n/n.lag), %>% mutate(n.lag = n.lag + 0.00001, n = n + 0.00001, way to bring S?
- [x] center or scale with preProcess?? preProc = c("center", "scale") makes no difference
- [ ] direction of relationship, or cor()
- [x] importance over time
- [x] spatial sampling
- [ ] hierarchy: ranking correlation and line graphs 
- [ ] OA

@@

for (t in years){
  tplus1 <- t + 1
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
                     data = df %>% filter(year==t | year==tplus1), #%>% 
                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
                     trControl = tc,
                     method = "rf", #"rpart", # rf lm
                     na.action = na.omit,
                     preProc = c("center", "scale"),
                     importance = TRUE)
}

@@

## RF spatial

```{r}
df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
  ungroup() %>% 
  dplyr::select(-help)

df <- df %>% dplyr::select(-n.lag)

### train

```{r}
#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   #seeds = seeds,
                   allowParallel = T,
                   savePredictions = 'final')

# run the model for every 2 years
start.time <- Sys.time()
regions <- df %>% distinct(RGN21CD)
regions <- as.vector(regions$RGN21CD)

for (i in regions){
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
                     data = df %>% filter(RGN21CD==i), trControl = tc,
                     method = "xgbTree", # rf "rpart", # rf lm
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = TRUE)
  assign(paste0("model.all",as.character(i)), model.all)
}

```{r}
# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

# use the model trainned in year t to predict io of year t + 1
#pred.by.year.all <- data.frame(matrix(NA,nrow=336, ncol = 0))
for (i in regions){
  for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", i, ".on.", j), pred)
    #pred.by.year.all <- rbind(pred.by.year.all, pred)
  }
}

df.predictions <- data.frame()
for (i in regions){
  for (j in regions[!regions == i]){
    d <- paste0("region.predict.model.", i, ".on.", j)
    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
    d <- d[,c(5, 2, 4, 6, 1)]
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  }
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))

# pred.by.year to list by year
pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11

# calculate metrics for every year
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth
rf.year.all.metrics %>% as.data.frame() %>%
  round(2) %>%
  kable()

# there are 2 columns named year, so I drop one
pred.by.region.all$year <- NULL

ggplot(data = df.predictions, aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ test.train, ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```
